Developing a preprint culture in biology

### 2012-10-05 ms.tex
minor

% I added the following three paragraphs to the conclusion, offering also a
% discussion on the on going explosion of different publication models. Take
% what you want and remove what you don't need.  I also had the feeling we might
% want to include a single paragraph, very short, calling that evaluations
% (academia, grant agencies etc...) looking at publication records or report of
% scientific activities start to view preprints as a significant and sound
% scientific contribution. One of the most important benefit is again the speed
% of the publication process. Most grants are for three years or less. It makes
% it hard to have a published paper within the limits of the grant. Preprints
% could however speed the process and show that contributions were made during
% the time of the grant. 
The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers, wether
they are junior looking for an academic position or senior trying to secure
funding for their team, to orient their publication strategy in order to
maximize their number of publications and total citations. A well-known
consequence is to submit manuscripts first to the most prestigious journals, and
then resubmit to lower level journals as they are rejected. The numerous
negative impacts of such behavior have been discussed in depth \cite{hoc09}.
They all contribute somehow in a general slowing down of scientific progress.
Research activities and the publication process are drifting away from their
fundamental object, namely the diffusion of novel scientific discoveries.

@@

The current opening of the publication process we are currently witnessing with
the arrival of new preprint systems and innovative publication models such as
F1000, PLoS One, Nature Scientific Reports, might significantly bypass the
inflation of bibliographic metrics. They not only promote scientific community
by changing the focus of the peer-reviewing, they also favor the publication of
results that might otherwise be rejected because of low interest, negative
results, descriptive and quantitative studies and datasets. One of drawback of
making publication easier is obviously the proliferation of studies of uneven
quality. A tradeoff between the intensity of the peer-review filtering and the
benefits to science has been hypothesized \cite{Aarssen2012}. With increasingly
stringent peer reviewing, the quality of published papers and easiness of
finding discoveries might increase, at the cost of censorship, increasing load
on authors and reviewers and time for publication. Preprints are simply
bypassing this model, for what we believe is the progress of science: they speed
up the dissemination of scientific discoveries, impede censorship and put on
reader's shoulders the responsibility to judge originality and pertinence.

>>>>

% Dominique:
% I added the following three paragraphs to the conclusion, offering also a
% discussion on the on going explosion of different publication models. Take
% what you want and remove what you don't need.  I also had the feeling we might
% want to include a single paragraph, very short, calling that evaluations
% (academia, grant agencies etc...) looking at publication records or report of
% scientific activities start to view preprints as a significant and sound
% scientific contribution. One of the most important benefit is again the speed
% of the publication process. Most grants are for three years or less. It makes
% it hard to have a published paper within the limits of the grant. Preprints
% could however speed the process and show that contributions were made during
% the time of the grant.

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers, wether
they are junior looking for an academic position or senior trying to secure
funding for their team, to orient their publication strategy in order to
maximize their number of publications and total citations. A well-known
consequence is to submit manuscripts first to the most prestigious journals, and
then resubmit to lower level journals as they are rejected. The numerous
negative impacts of such behavior have been discussed in depth \cite{hoc09} and
include a long delay between the time a manuscript is finished to its
publication.  They all contribute somehow in a general slowing down of
scientific progress.  Research activities and the publication process are
drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

@@

The current opening of the publication process we are currently witnessing with
the arrival of new preprint systems and innovative publication models such as
F1000, PLOS One, Nature Scientific Reports, might significantly bypass the
inflation of bibliographic metrics. They not only promote scientific community
by changing the focus of the peer-reviewing, they also favor the publication of
results that might otherwise be rejected because of low interest, negative
results, descriptive and quantitative studies and datasets. One of drawback of
making publication easier is obviously the proliferation of studies of uneven
quality. A tradeoff between the intensity of the peer-review filtering and the
benefits to science has been hypothesized \cite{Aarssen2012}. With increasingly
stringent peer reviewing, the quality of published papers and easiness of
finding discoveries might increase, at the cost of censorship, increasing load
on authors and reviewers and time for publication. Preprints are simply
bypassing this model, for what we believe is the progress of science: they speed
up the dissemination of scientific discoveries, impede censorship and put on
reader's shoulders the responsibility to judge originality and pertinence.
